What's in a Name? Linear Temporal Logic Literally Represents Time Lines

### 2023-06-18 paper/main.tex
something
update figure 11
update paper

\documentclass[preprint,12pt]{elsarticle}
\journal{Journal of Computer Languages}
\usepackage[x11names, svgnames, rgb]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm,tikz, subfig}
\usetikzlibrary{positioning,shapes,quotes,calc,snakes,arrows}

@@

\begin{frontmatter}
    \title{What's in a Name? Linear Temporal Logic Literally Represents Time Lines}

	\author[inst1]{Runming Li\corref{cor1}}
    \ead{runmingl@andrew.cmu.edu}

	\author[inst1]{Keerthana Gurushankar\corref{cor1}}
    \ead{kgurusha@andrew.cmu.edu}

    \author[inst1]{Marijn~J.H.~Heule\corref{cor2}}
    \ead{marijn@cmu.edu}

    \author[inst2]{Kristin~Y.~Rozier}
    \ead{kyrozier@iastate.edu}

    \affiliation[inst1]{organization={Carnegie Mellon University},%Department and Organization
                %addressline={5000 Forbes Ave},
                city={Pittsburgh},
                %postcode={15213},
                state={PA},
                country={USA}}

    \affiliation[inst2]{organization={Iowa State University},%Department and Organization
                %addressline={Address Two},
                city={Ames},
                %postcode={50010},
                state={IA},
                country={USA}}

	\cortext[cor1]{These authors contributed equally to this work}
    \cortext[cor2]{Corresponding author}

    \begin{abstract}
        Linear Temporal Logic (\ltl) is widely used to specify requirements in safety-critical systems.
        However, like many formal verification techniques, it is known to be unintuitive and error-prone for human practitioners to specify and validate.
        In this paper, we provide a new timeline tool for visualizing \ltl-based specifications, which is effective at intuitively representing a wide range of formulas.
        Our tool generates timeline visualizations by translating \ltl\ formulae to intermediate representations as \Buchi automata and then $\omega$-regular expressions, and finally simplifying and visualizing the expressions.
        We provide an algorithm for this visualization, a theoretical soundness analysis, and an implementation.
    \end{abstract}

    \begin{highlights}
        \item We transform Linear Temporal Logic (\ltl) formulas to timeline visualizations for the set of satisfying traces.
        \item We translate \ltl\ formulas to \Buchi automata, then to $\omega$-regular expressions, then simplify them and use graph visualization tools to draw timelines.
        \item Our algorithm is sound and complete.
        \item Our tool, \tool, effectively constructs small timeline visualizations for a range of \ltl\ formulas.
    \end{highlights}

    \begin{keyword}
        Modal and Temporal Logics
        \sep Logic and Verification
        \sep Regular languages
    \end{keyword}
\end{frontmatter}
\section{Introduction}

Requirement specification is a central step in the development of safety-critical systems. As a first step, requirements are typically written in natural language. For example, here is a real-world requirement specification from an air traffic control system ~\cite{ZR14}
\begin{center}
    ``If a TSAFE command is sent to an aircraft, controller/AutoResolver should then hand off the control of this aircraft''
\end{center}
Such natural language requirements are often ambiguous and not amenable to formal analysis, so requirements must often be specified using rigorous formalized semantics for the purpose of verification and validation. The above requirement, when translated to Linear Temporal Logic (detailed in \cref{sec:ltl}), may be expressed as follows:
\begin{align*}
    \always (\texttt{tsafe.TSAFE\_command1} \land \texttt{controller.CTR\_control\_1} \\
    \limplies \nextt (\neg \texttt{controller.CTR\_control\_1}))
\end{align*}
Several tools have been developed to reason about formal specifications written with such \ltl\ formulas, like R2U2 and FRET \cite{GPMS20}. However, these formalized semantics are unintuitive and error-prone for human practitioners, and the tools are much less useful when system designers can't validate input specifications.
Thus, a significant hurdle remains: how can we convincingly demonstrate to the humans in the loop, from system designers to certifiers, that the analyzed formulas truly represent the desired system requirements? % quote MLTL paper?

@@

Our tool, \tool, uses a remarkably simple algorithm: transforming an input \ltl\ formula through a sequence of automaton and regular expression-based intermediate representations, and is surprisingly effective at providing small representations for a range of \ltl\ formulas. It would visualize the above air traffic control requirement behavior with the timeline shown in \cref{example:air}.

@@

\subsection{Linear temporal logic (LTL)} \label{sec:ltl}

\begin{definition}[Linear Temporal Logic (LTL)]
    % cite Kristin's survey paper
    The syntax of an \ltl\ formula over a set of atomic propositions $\AP$, where $p\in\AP$ is a propositional variable, is given by the following grammar:
    \begin{align*}
        \Phi ::= p \mid \neg \Phi \mid \Phi \land \Phi \mid \Phi \lor \Phi \mid \Phi \limplies \Phi \mid \always \Phi \mid \eventually \Phi \mid \nextt \Phi \mid \Phi \stronguntil \Phi \mid \Phi \weakrelease \Phi
    \end{align*}\label{ltl-defn}
\end{definition}

@@

\begin{definition}[Semantics of \ltl\ formula]
    The semantics of \ltl\ formula is defined via judgment $\pi, i \models \Phi$ where $\pi : \omega \rightarrow 2^{\AP}$ is the computation that stores the truthhood and falsehood of every atomic proposition at every time step, and $\omega$ is the set of natural numbers that denote the time step. The semantics follows exactly from Rozier~\cite{Roz11}.
\end{definition}

\subsection{State-based \Buchi automata (BA)}
\begin{definition}[$\omega$-word]
    An $\omega$-word or infinite run of $\Sigma$, is an infinite string $s = (s_0, s_1, s_2, \dots)$ where each $s_i\in \Sigma$
\end{definition}
\begin{definition}[\Buchi automaton]
    A \Buchi automaton is a $5$-tuple, $G = (Q, \Sigma, \delta, s, F)$ consisting of
    \begin{enumerate}
        \item a finite set of states $Q$
        \item a finite alphabet of input symbols called the alphabet $\Sigma$
        \item a partial function $\delta : Q\times \Sigma \to Q$, called the transition function
        \item an initial or start state called $s\in Q$
        \item a set of accept states $F \subseteq Q$
    \end{enumerate}
    $G$ accepts an infinite run iff at least one of its infinitely visited states is in $F$
\end{definition}

\subsection{$\omega$-regular expression}
\begin{definition}[Regular expression]
    For word $a \in \Sigma$, regular expression is defined by the following grammar:
    \begin{align*}
        A ::= \emptyset \mid \epsilon \mid a \mid AA \mid A + A \mid A^*
    \end{align*}
\end{definition}
\begin{definition}[Semantics of regular expression]
    Let $L(A)$ denote the language accepted by regular expression $A$. Then $L(A)$ is inductively defined as
    \begin{align*}
        L(\emptyset) & = \emptyset \\
        L(\epsilon) & = \{\epsilon\} \tag{$\epsilon$ denotes the empty string}\\
        L(a) & = \{a\} \\
        L(A_1A_2) & = \{s_1s_2 \mid s_1 \in L(A_1) \text{ and } s_2 \in L(A_2)\} \\
        L(A_1 + A_2) & = L(A_1) \cup L(A_2) \\
        L^{(0)}(A) & = \{\epsilon\} \\
        L^{(i + 1)}(A) & = \{s_1s_2 \mid s_1 \in L(A) \text{ and } s_2 \in L^{(i)}(A)\} \\\
        L(A^*) & = \bigcup_{i \ge 0} L^{(i)}(A)
    \end{align*}
\end{definition}

\begin{remark}
    For the purpose of our tool, $\Sigma$ is the set of propositional logic formula defined as
    \[
    \Psi ::= p \mid \neg \Psi \mid \Psi \land \Psi \mid \Psi \lor \Psi \mid \Psi \limplies \Psi
    \]
\end{remark}

\begin{definition}[$\omega$-regular expression]
    Regular expression concerns only finite-length strings. However, since \ltl\ formulas reason about events that happen over an infinite-length timeline, we need to model it using infinite regular expressions (i.e. $\omega$-regular expression). $\omega$-regular expression is defined by the following grammar:
    \begin{align*}
        B ::= A^{\omega} \mid AB \mid B + B
    \end{align*}
\end{definition}
\begin{definition}[Semantics of $\omega$-regular expression]\label{def:omega-semantics}
    Let $\Sigma^{\omega}$ denote the set of infinite-length string over fixed alphabet $\Sigma$. Let $L_{\omega}(B)$ denote the $\omega$-language accepted by $\omega$-regular expression $B$. Then $L_{\omega}(B)$ is inductively defined as
    \begin{align*}
        L_{\omega}(A^{\omega}) & = \{s_1s_2s_3\cdots \mid s_i \in L(A) \text{ and } i \ge 1\} \\
        L_{\omega}(AB) & = \{s_1s_2 \mid s_1 \in L(A) \text{ and } s_2 \in L_{\omega}(B)\} \\
        L_{\omega}(B_1 + B_2) &= L_{\omega}(B_1) \cup L_{\omega}(B_2)
    \end{align*}
\end{definition}

\subsection{Timeline}
We present timelines as graphic visualizations. It contains the following features:
\begin{itemize}
    \item Every timeline starts with a node named ``start''.
    \item Every node represents one time step, and each node has a propositional logic formula $\Psi$, which specifies the behavior of atomic propositions at that time step. The formula $\Psi$ must be true at that time step. If $\Psi = 1$, that means all atomic propositions can behave arbitrarily.
    \item A node in egg shake with a caption ``repeats $0$ - $\infty$'' means to repeat the (one time step) current node for arbitrarily finitely many times.
    \item A node with label ``$\cdots$'' means to repeat the pattern prior to it and after it arbitrarily finitely many times.
    \item The grey box means repeat infinitely. Once we reach the end of a timeline in the grey box, we must reenter the same grey box from any of its starting point.
\end{itemize}

\begin{example}
    In \cref{fig:ex2}, one can reason about two parallel timelines: the upper timeline starts with $p$ at the first time step, followed by entering the grey box with one step of $\lnot p$ and one step of $p$. Then at the end of the grey box, we reenter the box, with the next time step being $\lnot p$, and so on. The lower timeline starts with one step of $\lnot p$ and one step of $p$ outside the grey box, and then enter the infinite run of $\lnot p$ and $p$ repeating.
    \begin{figure}[!h]
        \centering
        \begin{tikzpicture}
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p1) {\!\!\!$p$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p2) [below=0.4cm of p1] {\!\!\!$p$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p3) [left= 1cm of p2] {\!\!\!$\lnot p$\!\!\!};
        \node (m) at ($(p1)!0.5!(p2)$) {};
        \node[shape=rectangle, fill=gray!30!white, style={minimum width=4.3cm, minimum height=1.25cm}] (g) [right= 1.35cm of m] {};

@@

@@

\section{Algorithm}
On a high level, our algorithm of converting \ltl\ formulas into timeline visualization works as follows:
\begin{enumerate}
    \item Convert the given \ltl\ formula to its corresponding \Buchi automata % (as detailed in \cref{ltl2aut})
    \item Derive the $\omega$-regular expression corresponding to the \Buchi automata % (in \cref{aut2regex})
    \item Simplify the derived $\omega$-regular expression % (in \cref{regex-simplify})
    \item Visualize the $\omega$-regular expression according to its structure %(in \cref{regex2timeline})
\end{enumerate}
\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{img/algorithm_outline.pdf}
    \caption{Algorithm outline}
    \label{fig:algo}
\end{figure}

@@

The $\omega$-regular expression generated in \cref{aut2regex} may not be the ``simplest'' for the purpose of visualizing the timeline. We have observed multiple patterns in the resulting $\omega$-regular expression that could be simplified. For example, an $\omega$-regular expression in the form of $r^*r^{\omega}$ represents the same timeline as $r^{\omega}$, but the latter is more intuitive. For this purpose, we devised some simplification rules in our tool, based on our observation of common patterns in the generated $\omega$-regular expression.

\paragraph*{Rule-based simplification}
Here we show a demonstrating subset of simplification rules we encoded.
\begin{alignat*}{2}
        & r_1 + r_1r_2^* && \Longrightarrow r_1r_2^* \\
        & r + r && \Longrightarrow r \\
        & r_1 + r_2^*r_1 && \Longrightarrow r_2^*r_1 \\
        & (r^*)^{\omega} && \Longrightarrow r^{\omega} \\
        & (r_1r_2^*)r_2^{\omega} && \Longrightarrow r_1r_2^{\omega} \\
        & (r_1r_2)r_2^{\omega} && \Longrightarrow r_1r_2^{\omega} \\
        & r^*r^{\omega} && \Longrightarrow r^{\omega} \\
        & rr^{\omega} && \Longrightarrow r^{\omega}
\end{alignat*}

@@

    The formula $\Phi = \always (a \limplies \eventually (\neg a))$ generates un-simplified $\omega$-regular expression $((\lnot a) | (aa^*(\lnot a)))((\lnot a) | (aa^*(\lnot a)))^{\omega}$, and simplified version $((\lnot a) | (aa^*(\lnot a)))^{\omega}$, which corresponds to the two timelines in \cref{fig:unsimplified} and \cref{fig:simplified}.
    \begin{figure}[h!]

@@

        \caption{Timeline for %un-simplified

@@

    \begin{figure}[h!]

@@

        \caption{Timeline for simplified $\omega$-regular expression $((\lnot a) | (aa^*(\lnot a)))^{\omega}$}

@@

@@

\subsection{$\omega$-regex to timeline} \label{regex2timeline}
Our tool uses Graphviz~\cite{Ellson2001GraphvizO} to achieve the timeline visualization step. By construction of our algorithm, every $\omega$-regular expression is in the form of
\[
    A_1A_2^{\omega} + A_3A_4^{\omega} + \cdots + A_{2n-1}A_{2n}^{\omega}
\]
where $A_i$ are regular expressions, and $A_{2i-1}$ could be $\epsilon$. On a high level, for each of regular expression $A_i$, we visualize it as a set of accepted input. We view each union operator as parallel timelines, and each $A_{2i-1}$ is concatenated with $A_{2i}$ repeated infinitely many times, as represented in the grey box. \cref{fig:timeline} presents a generic timeline resulting from this form of construction.
\begin{figure}[h!]
    \centering
    \begin{tikzpicture}
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {$A_1$};
    \node[shape=rectangle, fill=gray!30!white, style={minimum width=1.8cm, minimum height=1.3cm}] (a2g) [right= 0.85cm of a1] {};
    \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [right= 1cm of a1] {$A_2$};
    \node[draw, thick, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a3) [below= 0.4cm of a1] {$A_3$};
    \node[shape=rectangle, fill=gray!30!white, style={minimum width=1.8cm, minimum height=1.3cm}] (a4g) [right= 0.85cm of a3] {};
    \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a4) [right= 1cm of a3] {$A_4$};
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a5) [below= 0.4cm of a3] {...};
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a7) [below= 0.4cm of a5] {\!\!\!$A_{2n-1}$\!\!\!};
    \node[shape=rectangle, fill=gray!30!white, style={minimum width=1.8cm, minimum height=1.3cm}] (a8g) [right= 0.85cm of a7] {};
    \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a8) [right= 1cm of a7] {\!\!\!$A_{2n}$\!\!\!};
    \node (m) at ($(a1)!0.5!(a7)$) {};
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) [left= 1.5cm of m] {\!\!\!\!start\!\!\!\!};
    \draw[thick] (s) edge[-latex] (a1) (s) edge[-latex] (a3) (s) edge[-latex] (a5) (s) edge[-latex] (a7);
    \draw[thick] (a1) edge[-latex] (a2) (a3) edge[-latex] (a4) (a7) edge[-latex] (a8);
    \end{tikzpicture}
%    \includegraphics[scale=0.5]{img/timeline.png}
    \caption{Generic timeline construction from $A_1A_2^{\omega} + A_3A_4^{\omega} + \cdots + A_{2n-1}A_{2n}^{\omega}$}
    \label{fig:timeline}
\end{figure}

@@

        \always (\texttt{tsafe.TSAFE\_command1} \land \texttt{controller.CTR\_control\_1} \\
        \limplies \nextt (\neg \texttt{controller.CTR\_control\_1}))

@@

    \begin{figure}[h!]

@@

>>>>

\documentclass[conference]{IEEEtran}
\usepackage[x11names, svgnames, rgb]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm,tikz, subfig}
\usetikzlibrary{positioning,shapes,quotes,calc,snakes,arrows}

@@

\title{What's in a Name? Linear Temporal Logic Literally Represents Time Lines}

\author{\IEEEauthorblockN{Runming Li}
\IEEEauthorblockA{\textit{Carnegie Mellon University}\\
Pittsburgh, USA \\
runmingl@andrew.cmu.edu}
\and
\IEEEauthorblockN{Keerthana Gurushankar}
\IEEEauthorblockA{\textit{Carnegie Mellon University}\\
Pittsburgh, USA \\
kgurusha@andrew.cmu.edu}
\and
\IEEEauthorblockN{Marijn~J.H.~Heule}
\IEEEauthorblockA{\textit{Carnegie Mellon University}\\
Pittsburgh, USA \\
marijn@cmu.edu}
\and
\IEEEauthorblockN{Kristin~Y.~Rozier}
\IEEEauthorblockA{\textit{Iowa State University} \\
Ames, USA \\
ekyrozier@iastate.edu}
}

\maketitle

\thispagestyle{plain}
\pagestyle{plain}

\begin{abstract}
  Linear Temporal Logic (LTL) is widely used to specify requirements in safety-critical systems.
  However, like many formal verification techniques, it is known to be unintuitive and error-prone for human practitioners to specify and validate.
  In this paper, we provide a new timeline tool for visualizing LTL-based specifications, which is effective at intuitively representing a wide range of formulas.
  Our tool generates timeline visualizations by translating LTL formulas to intermediate representations as B\"{u}chi automata and then regular expressions, and finally simplifying and visualizing the expressions.
  We provide an algorithm for this visualization, a theoretical soundness analysis, and an implementation.
\end{abstract}

\begin{IEEEkeywords}
Modal and Temporal Logics, Logic and Verification, Regular languages
\end{IEEEkeywords}

\section{Introduction}
Requirement specification is a central step in the development of safety-critical systems. As a first step, requirements are typically written in natural language. For example, here is a real-world requirement specification from an air traffic control system ~\cite{ZR14}
\begin{center}
    ``If a TSAFE command is sent to an aircraft, controller/AutoResolver should then hand off the control of this aircraft''
\end{center}
Such natural language requirements are often ambiguous and not amenable to formal analysis, so requirements must often be specified using rigorous formalized semantics for the purpose of verification and validation. The above requirement, when translated to Linear Temporal Logic (detailed in \cref{sec:ltl}), may be expressed as follows:
\begin{align*}
    \always & (\texttt{tsafe.TSAFE\_command1} \land \\ & \texttt{controller.CTR\_control\_1} \\
    \limplies & \nextt (\neg \texttt{controller.CTR\_control\_1}))
\end{align*}
Several tools have been developed to reason about formal specifications written with such \ltl\ formulas, like R2U2 and FRET \cite{GPMS20}. However, these formalized semantics are unintuitive and error-prone for human practitioners, and the tools are much less useful when system designers can't validate input specifications.
Thus, a significant hurdle remains: how can we convincingly demonstrate to the humans in the loop, from system designers to certifiers, that the analyzed formulas truly represent the desired system requirements? % quote MLTL paper?

@@

Our tool, \tool, uses a remarkably simple algorithm: transforming an input \ltl\ formula through a sequence of automata and regular expression-based intermediate representations, and is surprisingly effective at providing small representations for a range of \ltl\ formulas. It would visualize the above air traffic control requirement behavior with the timeline shown in \cref{example:air}.

@@

\subsection{Linear temporal logic (\ltl)} \label{sec:ltl}

\begin{definition}[Linear Temporal Logic (\ltl)]
    % cite Kristin's survey paper
    The syntax of an \ltl\ formula over a set of atomic propositions $\AP$, where $p\in\AP$ is a propositional variable, is given by the following grammar:
    \begin{align*}
        \Phi ::= & p \mid \neg \Phi \mid \Phi \land \Phi \mid \Phi \lor \Phi \mid \Phi \limplies \Phi \mid \always \Phi \mid \eventually \Phi \mid \nextt \Phi \\ & \mid \Phi \stronguntil \Phi \mid \Phi \weakrelease \Phi
    \end{align*}\label{ltl-defn}
\end{definition}

@@

\begin{definition}[Semantics of \ltl\ formula]
    The semantics of \ltl\ formula is defined via  $\pi, i \models \Phi$ where $\pi : \omega \rightarrow 2^{\AP}$ is the computation that stores the truthhood and falsehood of every atomic proposition at every time step, and $\omega$ is the set of natural numbers that denote the time step. The semantics follows exactly from Rozier~\cite{Roz11}.
\end{definition}

\subsection{State-based \Buchi automata (BA)}
\begin{definition}[$\omega$-word]
    An $\omega$-word or infinite run of $\Sigma$, is an infinite string $s = (s_0, s_1, s_2, \dots)$ where each $s_i\in \Sigma$
\end{definition}
\begin{definition}[\Buchi automaton]
    A \Buchi automaton is a $5$-tuple, $G = (Q, \Sigma, \delta, s, F)$ consisting of
    \begin{enumerate}
        \item a finite set of states $Q$
        \item a finite alphabet of input symbols called the alphabet $\Sigma$
        \item a partial function $\delta : Q\times \Sigma \to Q$, called the transition function
        \item an initial or start state called $s\in Q$
        \item a set of accepting states $F \subseteq Q$
    \end{enumerate}
    $G$ accepts an infinite run iff at least one of its infinitely visited states is in $F$
\end{definition}

\subsection{$\omega$-regular expression}
\begin{definition}[Regular expression]\label{def:regex}
    For word $a \in \Sigma$, regular expression is defined by the following grammar:
    \begin{align*}
        A ::= \emptyset \mid \epsilon \mid a \mid AA \mid A + A \mid A^*
    \end{align*}
\end{definition}
\begin{definition}[Semantics of regular expression]
    Let $L(A)$ denote the language accepted by regular expression $A$. Then $L(A)$ is inductively defined as
    \begin{align*}
        L(\emptyset) & = \emptyset \\
        L(\epsilon) & = \{\epsilon\} \tag{$\epsilon$ denotes the empty string}\\
        L(a) & = \{a\} \\
        L(A_1A_2) & = \{s_1s_2 \mid s_1 \in L(A_1) \text{ and } s_2 \in L(A_2)\} \\
        L(A_1 + A_2) & = L(A_1) \cup L(A_2) \\
        L^{(0)}(A) & = \{\epsilon\} \\
        L^{(i + 1)}(A) & = \{s_1s_2 \mid s_1 \in L(A) \text{ and } s_2 \in L^{(i)}(A)\} \\\
        L(A^*) & = \bigcup_{i \ge 0} L^{(i)}(A)
    \end{align*}
\end{definition}

\begin{remark}
    For the purpose of our tool, $\Sigma$ is the set of propositional logic formulas defined as
    \[
    \Psi ::= p \mid \top \mid \bot \mid \neg \Psi \mid \Psi \land \Psi \mid \Psi \lor \Psi \mid \Psi \limplies \Psi
    \]
\end{remark}

\begin{definition}[$\omega$-regular expression]\label{def:omega-regex}
    Regular expression concerns only finite-length strings. However, since \ltl\ formulas reason about events that happen over an infinite-length timeline, we need to model it using infinite regular expressions (i.e. $\omega$-regular expression). $\omega$-regular expression is defined by the following grammar:
    \begin{align*}
        B ::= A^{\omega} \mid AB \mid B + B
    \end{align*}
\end{definition}
\begin{definition}[Semantics of $\omega$-regular expression]\label{def:omega-semantics}
    Let $\Sigma^{\omega}$ denote the set of infinite-length string over fixed alphabet $\Sigma$. Let $L_{\omega}(B)$ denote the $\omega$-language accepted by $\omega$-regular expression $B$. Then $L_{\omega}(B)$ is inductively defined as
    \begin{align*}
        L_{\omega}(A^{\omega}) & = \{s_1s_2s_3\cdots \mid s_i \in L(A) \text{ and } i \ge 1\} \tag{$\epsilon
         \not\in L(A)$} \\
        L_{\omega}(AB) & = \{s_1s_2 \mid s_1 \in L(A) \text{ and } s_2 \in L_{\omega}(B)\} \\
        L_{\omega}(B_1 + B_2) &= L_{\omega}(B_1) \cup L_{\omega}(B_2)
    \end{align*}
\end{definition}

\begin{remark}
  \Cref{def:regex,def:omega-regex} are the standard definitions of regular expressions and $\omega$-regular expressions, respectively. We include them here for completeness.
\end{remark}

\subsection{Timeline}\label{sec:timeline}
We present timelines as graphic visualizations. It contains the following features:
\begin{itemize}
    \item Every timeline starts with a node named ``start''.
    \item Every node represents one time step, and each node has a propositional logic formula $\Psi$, which specifies the behavior of atomic propositions at that time step. The formula $\Psi$ must be true at that time step. If $\Psi = 1$, that means all atomic propositions can behave arbitrarily.
    \item A node in egg shape with a caption ``repeats $0$ - $\infty$'' means to repeat the (one time step) current node for arbitrarily finitely many times.
    \item A node with label ``$\cdots$'' means to repeat the pattern prior to it and after it arbitrarily finitely many times.
    \item The grey box means repeat infinitely. Once we reach the end of a timeline in the grey box, we must reenter the same grey box from any of its (left) starting point.
    \item Timelines can be in parallel, signifying that either of the parallel timeline could happen.
\end{itemize}

The definition of timeline we use here are specifically what we devised for the purpose of visualizing \ltl\ formulas. We show how to construct such timelines from \ltl\ formulas in \cref{sec:correctness-timeline}.

\begin{example}
    In \cref{fig:ex2}, one can reason about two parallel timelines: the upper timeline starts with $p$ at the first time step, followed by entering the grey box with one step of $\lnot p$ and one step of $p$. Then at the end of the grey box, we reenter the box, with the next time step being $\lnot p$, and so on. The lower timeline starts with one step of $\lnot p$ and one step of $p$ outside the grey box, and then enter the infinite run of $\lnot p$ and $p$ repeating.
    \begin{figure}[!h]
        \centering
        \begin{tikzpicture}[scale=0.7, transform shape]
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p1) {\!\!\!$p$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p2) [below=0.4cm of p1] {\!\!\!$p$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p3) [left= 1cm of p2] {\!\!\!$\lnot p$\!\!\!};
        \node (m) at ($(p1)!0.5!(p2)$) {};
        \node[shape=rectangle, fill=gray!30!white, style={minimum width=4.3cm, minimum height=1.25cm}] (g) [right= 1.35cm of m] {};

@@

        \begin{tikzpicture}[>=latex',line join=bevel, scale=0.7, transform shape]

@@

\section{Algorithm}
On a high level, our algorithm of converting \ltl\ formulas into timeline visualizations works as follows:
\begin{enumerate}
    \item Convert the given \ltl\ formula to its corresponding \Buchi automaton. % (as detailed in \cref{ltl2aut})
    \item Derive the $\omega$-regular expression corresponding to the \Buchi automaton. % (in \cref{aut2regex})
    \item Simplify the derived $\omega$-regular expression. % (in \cref{regex-simplify})
    \item Visualize the $\omega$-regular expression according to its structure. %(in \cref{regex2timeline})
\end{enumerate}
\begin{figure}[!h]
    \centering
    % \includegraphics[width=\textwidth, scale=0.3]{img/algorithm_outline.pdf}
    TODO
    \caption{Algorithm outline}
    \label{fig:algo}
\end{figure}

@@

The $\omega$-regular expression generated in \cref{aut2regex} may not be the ``simplest'' for the purpose of visualizing the timeline. We have observed multiple patterns in the resulting $\omega$-regular expression that could be simplified. For example, an $\omega$-regular expression in the form of $r^*r^{\omega}$ represents the same timeline as $r^{\omega}$, but the latter is more intuitive and concise. For this purpose, we devised some simplification rules in our tool, based on our observation of common patterns in the generated $\omega$-regular expression.

There is no agreed upon canonical form for regular expression, hence we do not hope to find the simplest regular expression for visualization. Regular expression simplification comes down to regular expression equivalence checking, which is computationally hard \cite{MNS04}. For purpose of simplification, one could perform a search over equivalent regular expressions and decide which one is simpler to use. However, this strategy is expensive in terms of the running time, and for the purpose of visualization, we did not use this strategy in our tool. We consider this approach as future work.

\paragraph*{Rule-based simplification}
Here we show a demonstrating subset of simplification rules we encoded. In theory one could add more rules to the tool, so long as they are sound; here we only choose to encode the rules that represent common patterns we have observed in the generated $\omega$-regular expression.
\begin{alignat*}{2}
        & r_1 + r_1r_2^* && \Longrightarrow r_1r_2^* \\
        & r + r && \Longrightarrow r \\
        & r_1 + r_2^*r_1 && \Longrightarrow r_2^*r_1 \\
        & (r^*)^{\omega} && \Longrightarrow r^{\omega} \\
        & (r_1r_2^*)r_2^{\omega} && \Longrightarrow r_1r_2^{\omega} \\
        & (r_1r_2)r_2^{\omega} && \Longrightarrow r_1r_2^{\omega} \\
        & r^*r^{\omega} && \Longrightarrow r^{\omega} \\
        & rr^{\omega} && \Longrightarrow r^{\omega}
\end{alignat*}

@@

    Using our algorithm, the \ltl\ formula $\Phi = \always (a \limplies \eventually (\neg a))$ generates un-simplified $\omega$-regular expression $((\lnot a) | (aa^*(\lnot a)))((\lnot a) | (aa^*(\lnot a)))^{\omega}$, and simplified version $((\lnot a) | (aa^*(\lnot a)))^{\omega}$, which corresponds to the two timelines in \cref{fig:unsimplified} and \cref{fig:simplified}.
    \begin{figure*}[h!]

@@

        \caption{Timeline visualization for %un-simplified

@@

    \end{figure*}
    \begin{figure*}[h!]

@@

        \caption{Timeline visualization for simplified $\omega$-regular expression $((\lnot a) | (aa^*(\lnot a)))^{\omega}$}

@@

    \end{figure*}

@@

\subsection{$\omega$-regex to timeline} \label{regex2timeline}
Our tool uses Graphviz~\cite{Ellson2001GraphvizO} to achieve the timeline visualization step. By construction of our algorithm, every $\omega$-regular expression is in the form of
\[
    A_1A_2^{\omega} + A_3A_4^{\omega} + \cdots + A_{2n-1}A_{2n}^{\omega}
\]
where $A_i$ are regular expressions, $A_{2i-1}$ could be $\epsilon$, and $\epsilon \not\in L(A_{2i})$. On a high level, for each of regular expression $A_i$, we visualize it as a set of accepted input. We view each union operator as parallel timelines, and each $A_{2i-1}$ is concatenated with $A_{2i}$ repeated infinitely many times, as represented in the grey box. \cref{fig:timeline} presents a generic timeline resulting from this form of construction.
\begin{figure}[h!]
    \centering
    \begin{tikzpicture}
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {$A_1$};
    \node[shape=rectangle, fill=gray!30!white, style={minimum width=1.8cm, minimum height=1.3cm}] (a2g) [right= 0.85cm of a1] {};
    \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [right= 1cm of a1] {$A_2$};
    \node[draw, thick, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a3) [below= 0.4cm of a1] {$A_3$};
    \node[shape=rectangle, fill=gray!30!white, style={minimum width=1.8cm, minimum height=1.3cm}] (a4g) [right= 0.85cm of a3] {};
    \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a4) [right= 1cm of a3] {$A_4$};
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a5) [below= 0.4cm of a3] {...};
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a7) [below= 0.4cm of a5] {\!\!\!$A_{2n-1}$\!\!\!};
    \node[shape=rectangle, fill=gray!30!white, style={minimum width=1.8cm, minimum height=1.3cm}] (a8g) [right= 0.85cm of a7] {};
    \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a8) [right= 1cm of a7] {\!\!\!$A_{2n}$\!\!\!};
    \node (m) at ($(a1)!0.5!(a7)$) {};
    \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) [left= 1.5cm of m] {\!\!\!\!start\!\!\!\!};
    \draw[thick] (s) edge[-latex] (a1) (s) edge[-latex] (a3) (s) edge[-latex] (a5) (s) edge[-latex] (a7);
    \draw[thick] (a1) edge[-latex] (a2) (a3) edge[-latex] (a4) (a7) edge[-latex] (a8);
    \end{tikzpicture}
%    \includegraphics[scale=0.5]{img/timeline.png}
    \caption{Generic timeline construction from $A_1A_2^{\omega} + A_3A_4^{\omega} + \cdots + A_{2n-1}A_{2n}^{\omega}$}
    \label{fig:timeline}
\end{figure}

@@

        \always & (\texttt{tsafe.TSAFE\_command1} \land  \\ & \texttt{controller.CTR\_control\_1} \\
        \limplies & \nextt (\neg \texttt{controller.CTR\_control\_1}))

@@

    \begin{figure*}[h!]

@@

%        \begin{tikzpicture}
%            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) {\!\!\!\!start\!\!\!\!};
%            \node[shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab1) [right= of s]{\!\!\!\!$a \tand b$\!\!\!\!};
%            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab2) [above= 0.6cm of ab1]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
%            \node (m) at ($(ab1)!0.5!(ab2)$) {};
%            \node[shape=rectangle, fill=gray!30!white, style={minimum width=4cm, minimum height=3cm}] (g2) [right= -1cm of m] {};
